### 2012-02-03 paper_draft.tex
Fixed some typos, introduced JSON, changed some formulations

\subsubsection*{Connection to external services}
In order to provide users with relevant information on their respective genotypes, openSNP scans databases of the scientific literature for specific SNPs. A total number of 15,229 documents relevant to the SNP-IDs listed in openSNP could be found in the databases of Mendeley, the Public Library of Science and SNPedia. Of the primary literature, 25 \% are released in open access journals and can be accessed free of charge by every user (Figure n+2). For usability reasons, SNPs are ranked by the amount of information gathered by the external services.The external services themselves are ranked by how easily users can access information from these sources. The SNPedia entries are given the highest impact, as those are already manually curated, followed by open access publications out of the Public Library of Science. Lowest values are given to the Mendeley results, as those aren't necessary freely available to every user. An entry on SNPedia is valued 2.5 times as high as a PLoS publication and 5 times as high as a Mendeley entry.  
\subsubsection*{Data access}
OpenSNP offers extensive access to the data uploaded by users. Anyone can download single genotyping files for specific users, get archives of multiple genotyping files grouped by phenotypic variation, or access a single download that includes all genotyping files and all phenotypes in a comma-separated table. Additionally, the genetic data is accessible through the Distributed Annotation System, which offers all data for specific chromosomes and specific positions on single chromosomes. Users can discuss SNPs and phenotypes on the platform using a simple commenting system or private messages.

@@

By crowdsourcing the acquisition of genetic and phenotypic data, openSNP faces the same problems as any other open platform on the Internet, namely the need to trust users regarding the data they upload and enter on openSNP. Additionally, the quality of the data varies, especially in terms of accuracy on the phenotypic variation. While we try to suggest similar entries to the users, there are some cases where users wont follow those suggestions, so duplicates or similar phenotypes or varations in traits may arise.

@@

Another risk regarding data quality that should be kept in mind is a possible bias in data availability on openSNP, as we cannot rule out the possibility that only an unrepresentative subset of people buy DTC genetic testing and an even smaller subset is willing to publish the results, along with information of medical relevance.
\subsubsection*{Why should a user publish their genetic information? Pro vs. Con}
With openSNP, we have built a platform that can be used by customers of DTC genetic testing to easily share their genetic and phenotypic data with a wide audience, as well as by scientists and interested citizens who are looking for datasets to be used in GWAS. Customers of DTC genetic testing also benefit from an easy access to primary literature on SNPs and genetic variations they carry. While we don't have collected enough data to perform full scale GWAS yet, this might be possible in the future, as user numbers are rising. 
\subsubsection*{Privacy and health implications and ethical considerations}

@@

% You may title this section "Methods" or "Models". 
% "Models" is not a valid title for PLoS ONE authors. However, PLoS ONE
% authors may use "Analysis" 
\section*{Materials and Methods}
\subsection*{Survey on Sharing Genetic Information}
The survey was done with Google Docs and was distributed to possible participants through the 23andMe community, the DIYBiology mailing list, blogs which focus on genetics and DTC genetic testing and social media like Twitter, Google Plus and Facebook.

@@

Processes with a longer runtime, such as parsing the genotyping results, creating archives of results which are to be mailed to users and queries to external resources are handled using the ruby gem Resque and a Redis server. Search features on the platform itself are implemented using SOLR and the ruby gem Sunspot. Additionally, data can be requested from openSNP using the Distributed Annotation System \cite{Dowell2001,Jenkinson2008}. The required data is stored in a mySQL database, the delivery of the data is done by using ProServer \cite{Finn2007}, which was modified by Gel et al. for use in easyDAS \cite{GelMoreno2011}.

>>>>

\subsubsection*{Connection to external services}
In order to provide users with relevant information on their respective genotypes, openSNP scans databases of the scientific literature for specific SNPs. A total number of 15,229 documents relevant to the SNP-IDs listed in openSNP could be found in the databases of Mendeley, the Public Library of Science and SNPedia. Of the primary literature, 25 \% are released in open access journals and can be accessed free of charge by every user (Figure n+2). For usability reasons, SNPs are ranked by the amount of information gathered by the external services.The external services themselves are ranked by how easily users can access information from these sources. The SNPedia entries are given the highest impact, as those are already manually curated, followed by open access publications out of the Public Library of Science. Lowest values are given to the Mendeley results, as those aren't necessary freely available to every user. An entry on SNPedia is valued 2.5 times as high as a PLoS publication and 5 times as high as a Mendeley entry.

\subsubsection*{Data access}
OpenSNP offers extensive access to the data uploaded by users. Anyone can download single genotyping files for specific users, get archives of multiple genotyping files grouped by phenotypic variation, or access a single download that includes all genotyping files and all phenotypes in a comma-separated table. The genetic data is also accessible through the Distributed Annotation System, which offers all data for specific chromosomes and specific positions on single chromosomes. The data is additionally available over a JSON-API, which allows users to directly access data in the JSON-format. Users can discuss SNPs and phenotypes on the platform using a simple commenting system, or send each other private messages.

@@

By crowdsourcing the acquisition of genetic and phenotypic data, openSNP faces the same problems as any other open platform on the Internet, namely the need to trust users regarding the data they upload and enter on openSNP. Additionally, the quality of the data varies, especially in terms of accuracy on the phenotypic variation, with users entering data in different measurement systems. While we try to suggest similar entries to the users, there are some cases where users wont follow those suggestions, so duplicates or similar phenotypes or variations in traits may arise.

@@

Another risk regarding data quality that should be kept in mind is a possible bias in data availability on openSNP, as we cannot rule out the possibility that only an unrepresentative subset of people buy DTC genetic testing, from which an even smaller subset is willing to publish the results, along with phenotypic information of medical relevance.

\subsubsection*{Why should a user publish their genetic information? Pros vs. Cons}
With openSNP, we have built a platform that can be used by customers of DTC genetic testing to easily share their genetic and phenotypic data with a wide audience, as well as by scientists and interested citizens who are looking for datasets to be used in GWAS. Customers of DTC genetic testing also benefit from an easy access to primary literature on SNPs and genetic variations they carry. While we don't have collected enough data to perform full scale GWAS yet, this might be possible in the future, as user numbers are rising. 
\subsubsection*{Privacy and health implications and ethical considerations}

@@

% You may title this section "Methods" or "Models". 
% "Models" is not a valid title for PLoS ONE authors. However, PLoS ONE
% authors may use "Analysis" 
\section*{Materials and Methods}
\subsection*{Survey on Sharing Genetic Information}
The survey was performed using Google Docs and was distributed to possible participants through the 23andMe community, the DIYBiology mailing list, blogs which focus on genetics and DTC genetic testing and social media like Twitter, Google Plus and Facebook.

@@

Processes with a longer runtime, such as parsing the genotyping results, creating archives of results which are to be mailed to users and queries to external resources are handled using the ruby gem Resque and the standalone server Redis. Search features on the platform itself are implemented using SOLR and the ruby gem Sunspot. Additionally, data can be requested from openSNP using the Distributed Annotation System \cite{Dowell2001,Jenkinson2008}. The required data is stored in a PostgreSQL database, the delivery of the data is done by using ProServer \cite{Finn2007}, which was modified by Gel et al. for use in easyDAS \cite{GelMoreno2011}. % probably no proserver, more own stuff
